---
title: "basicStudyDesign"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# OVERVIEW
The purpose of this vignette is to demonstrate how to use the functions and data in the `import2ImmPort` package to
create the basicStudyDesign.csv that must be submitted along with assay and other metaData to the public ImmPort 
database at NIH.

First, we will load the `import2ImmPort` library, which has the functions and data we need.
```{r load-library}
library(Import2ImmPort)
```

Next, we need to create the 9 dataframes that will be put together to form the larger csv:
- study
- arm_or_cohort
- inclusion_exclusion
- planned_visit
- study_2_protocol
- study_file
- study_link
- study_personnel
- study_pubmed

The first dataframe, called *study* is unique in that it has two columns, one with names and the other with 
values.  In this way it functions more like a named list and is treated differently from most other dataframes 
that have a typical format with column headers and unlimited rows allowed.

In order to get a pre-made dataframe for *study* we use the `getTemplateDF()` function.  The only argument is the
name of the template.
```{r create-study-df}
study <- getTemplateDF("study")
DT::datatable(study)
```

Now that we have the *study* dataframe, the easiest way to edit the dataframe by hand is to use the `edit()` function
and save the output.  This function is found in the `utils` package, but depends on the `X11` library.  If you are on a 
newer mac you may need to install XQuartz from the project's website as it is no longer bundled.  If for some reason `edit()`
does not save the output correctly on your machine, then you will have to input information using an R-based approach.
```{r input-to-study, eval=FALSE}
edit(study)
```

```{r R-based-approach-to-editing}
study[ , 2 ] <- c("SdyID",
                  "myBriefTitle",
                  "myOfficialTitle",
                  "",
                  "This is a study",
                  "This is a longer description of the study",
                  "We hypothesize ... ",
                  "The objectives were ... ",
                  "Endpoints are ... ",
                  "NIH",
                  50,
                  "disease",
                  10,
                  30,
                  "years",
                  "doctor",
                  "01/01/2018")
```

Similar to *study*, the other 8 dataframes that make up the basicStudyDesign.csv can be accessed with the
`getTemplateDF()` function.  The next one we focus on is called *arm_or_cohort*.
```{r make-arm-or-cohort-df}
armOrCohort <- getTemplateDF("arm_or_cohort")
```



