---
title: "subjectsHuman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# OVERVIEW
The purpose of this vignette is to demonstrate how to use the functions and data in the 
`import2ImmPort` package to create the subjectHuman.csv that must be submitted along 
with assay and other metaData to the public ImmPort database at NIH for projects with patients. Animal subjects require a different template (subjectsAnimal).  

When creating your own submission documents, you can create a similar vignette so that the whole process is 
reproducible.

First, we will load the `import2ImmPort` library, which has the functions and templates we need.

```{r load-library}
library(Import2ImmPort)
```

We will need to create a data frame with the following variables:  
- User Defined ID  
- Arm Or Cohort ID  
- Protocol  
- Gender  
- Min Subject Age  
- Max Subject Age  
- Age Unit  
- Age Event  
- Age Event Specify  
- Subject Phenotype  
- Subject Location  
- Ancestry Populations  
- Ethnicity  
- Race  
- Race Specify  
- Description  
  
In order to get a pre-made dataframe for subjectsHuman we use the `getTemplateDF()` function.  
The only argument is the name of the template.

```{r create-df}
subjectsHuman <- getTemplateDF("subjectsHuman", 5)
DT::datatable(subjectsHuman)
```

Now that we have the *study* dataframe, the easiest way to edit the dataframe by hand is to 
use the `edit()` function and save the output by clicking "quit" in the editor.  This function 
is found in the `utils` package, but depends on the `X11` library.  If you are on a newer mac 
you may need to install XQuartz from the project's website as it is no longer bundled.  If for 
some reason `edit()` does not save the output correctly on your machine, then you will have to 
input information using an R-based approach.

```{r input-to-study, eval=FALSE}
str(subjectsHuman)
subjectsHuman$`Ancestry Population(s)` <- as.character(subjectsHuman$`Ancestry Population(s)`) ## Why is it read in like that?
edit(subjectsHuman)
```

An R-based approach example:

```{r R-based-approach-to-editing}
subjectsHuman$`User Defined ID` <- c(1:5)
subjectsHuman$`Arm Or Cohort ID` <-

DT::datatable(subjectsHuman)

```

