---
title: "subjectsHuman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# OVERVIEW
The purpose of this vignette is to demonstrate how to use the functions and data in the 
`import2ImmPort` package to create the subjectHuman.csv that must be submitted along 
with assay and other metaData to the public ImmPort database at NIH for projects with patients. Animal subjects require a different template (subjectsAnimal).  

When creating your own submission documents, you can create a similar vignette so that the whole process is 
reproducible.

First, we will load the `import2ImmPort` library, which has the functions and templates we need.

```{r load-library}
library(Import2ImmPort)
```

We will need to create a dataframe with the following variables:  
- User Defined ID  
- Arm Or Cohort ID  
- Protocol  
- Gender  
- Min Subject Age  
- Max Subject Age  
- Age Unit  
- Age Event  
- Age Event Specify  
- Subject Phenotype  
- Subject Location  
- Ancestry Populations  
- Ethnicity  
- Race  
- Race Specify  
- Description  
  
In order to get a pre-made dataframe for subjectsHuman we use the `getTemplateDF()` function.  
The only argument is the name of the template.

```{r create-df}
subjectsHuman <- getTemplateDF("subjectsHuman", 3)
DT::datatable(subjectsHuman)
```

Now that we have the *study* dataframe, the easiest way to edit the dataframe by hand is to 
use the `edit()` function and save the output by clicking "quit" in the editor.  This function 
is found in the `utils` package, but depends on the `X11` library.  If you are on a newer mac 
you may need to install XQuartz from the project's website as it is no longer bundled.  If for 
some reason `edit()` does not save the output correctly on your machine, then you will have to 
input information using an R-based approach.

```{r input-to-study, eval=FALSE}
edit(subjectsHuman)
```

An R-based approach example:

```{r R-based-approach-to-editing}
subjectsHuman$`User Defined ID` <- paste("myID", seq(1:3), sep = "_")
subjectsHuman$`Arm Or Cohort ID` <- paste("myArmAndCohortID", seq(1:3), sep = "_")
subjectsHuman$Gender <- c("Male", "Female", "Male")
subjectsHuman$`Min Subject Age` <- c(25, 22.3, 30)
subjectsHuman$`Max Subject Age` <- c(27, 24.3, 32)
subjectsHuman$`Age Unit` <- ""
subjectsHuman$`Age Event` <- "Age at Enrollment"
subjectsHuman$`Age Event Specify` <- "study milestone for subject's age that is unsupported by ImmPort"
subjectsHuman$`Subject Phenotype` <- "phenotype that captures subject disposition for the study"
subjectsHuman$`Subject Location` <- "United States of America"
subjectsHuman$`Ancestry Population(s)` <- "North America"
subjectsHuman$Ethnicity <- "Other"
subjectsHuman$Race <- "Asian"
subjectsHuman$`Race Specify` <- ""
subjectsHuman$Description <- ""

DT::datatable(subjectsHuman)

```

Now we can use the `checkTemplate()` function to ensure that our dataframe meets the ImmPort upload requirements. For the purpose of this vignette we will wrap the function in a `tryCatch()` method.  
```{r check-template}
results <- tryCatch(checkTemplate(subjectsHuman, "subjectsHuman"),
                    error = function(e){ return(e) }) 

# This will print out the error message
results$message
```

The error tells us that the required column `subjectsHuman$'Age Unit'` has been left blank. We must correct the dataframe and re-run the check function.
```{r fix-ageunit}
subjectsHuman$`Age Unit` <- "Years"

results <- tryCatch(checkTemplate(subjectsHuman, "subjectsHuman"),
                    error = function(e){ return(e) }) 

# This will print out the error message
results$message
```
Here the error message is telling us that a non-controlled term is being used in the `subjectsHuman$'Age Event'` column. To look up controlled terms we can use the `getLookupValues()` function. This function takes the template name and column name as arguments and returns a vector of the allowed terms.

```{r get-lookup}
getLookupValues("subjectsHuman", "Age Event")
```
From the output we can see that our original entry ('Age at Enrollment') must be corrected to 'Age at enrollment'. We can then check the dataframe again.

```{r correct-age-event}
subjectsHuman$`Age Event` <- "Age at enrollment"

checkTemplate(subjectsHuman, "subjectsHuman")
```

If there are no errors, the subjectsHuman dataframe is complete.  
  
------
  
### CREATING THE OUTPUT TSV

```{r transform-df}
transform_Data(dataName = "subjectsHuman",
               outputDir = "~/Documents/FHCRC/",
               validate = T)
```