---
title: "Create the Basic Study Design Template"
output: rmarkdown::html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Create Basic Study Design Template}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(DT.options = list(paginate = FALSE,
                          info = FALSE,
                          filter = FALSE))
```

### OVERVIEW
The purpose of this vignette is to demonstrate how to use the functions and data in the 
`R2i` package to create the basicStudyDesign.txt that must be submitted along 
with assay and other metaData to the public ImmPort database at NIH.  When creating your own
submission documents, you can create a similar vignette so that the whole process is 
reproducible.

### PREPARING THE STUDY DATA FRAME
First, we will load the `R2i` library, which has the functions and templates we need.

```{r load-library}
library(R2i)
```

Next, we need to create the 11 dataframes that will be put together to form the larger txt:

1. study
2. study_categorization
3. study_2_condition_or_disease
4. arm_or_cohort
5. study_personnel
6. planned_visit
7. inclusion_exclusion
8. study_2_protocol
9. study_file
10. study_link
11. study_pubmed

The first dataframe, called **study** is unique in that it has two columns, one with names 
and the other with values.  In this way it functions more like a named list and is treated 
differently from most other dataframes that have a typical format with column headers and 
unlimited rows allowed.

In order to get a pre-made dataframe for **study** we use the `getTemplateDF()` function.  
The only argument is the name of the template.

```{r create-study-df}
study <- getTemplateDF("study")
DT::datatable(study, options = list(scrollX = TRUE))
```

Now that we have the **study** dataframe, the easiest way to edit the dataframe by hand is to 
use the `edit()` function and save the output by clicking "quit" in the editor.  This function 
is found in the `utils` package, but depends on the `X11` library.  If you are on a newer mac 
you may need to install XQuartz from the project's website as it is no longer bundled.  If for 
some reason `edit()` does not save the output correctly on your machine, then you will have to 
input information using an R-based approach.

```{r input-to-study, eval=FALSE}
edit(study)
```

An R-based approach example:

```{r R-based-approach-to-editing}
study[1, ] <- list(NA,
                  "myBriefTitle",
                  "myOfficialTitle",
                  "This is a study",
                  "This is a longer description of the study",
                  "We hypothesize ... ",
                  "The objectives were ... ",
                  "Endpoints are ... ",
                  "NIH",
                  50,
                  10,
                  30,
                  "years",
                  "01/01/2018",
                  "01/01/2018"
                  )
```

We can check the **study** template to see if it passes the basic checks for class, dimension,
required columns, data types, and controlled terms by using `checkTemplate()`.  This function 
takes in the data frame we have created as well as the template name.  Any problems will throw 
an error, so for demonstration purposes we wrap the function in a `tryCatch()` method here.

```{r check-template-required-error}
# NOTE: messages are printed out
results <- tryCatch(checkTemplate(df = study),
                    error = function(e){ return(e) })

# to see error statement
results
```

It looks like we have an `NA` value for `study$'User Defined ID'`.  Let's correct this and
re-run our checks.

```{r correct-study-user-defined-id}
study$`User Defined ID` <- "sdyID"

# NOTE: messages are printed out
results <- tryCatch(checkTemplate(df = study),
                    error = function(e){ return(e) })

# to see error statement
results
```

It seems we are using a non-controlled term in `Age Unit` column, which requires a controlled value.  
Many templates use  controlled or preferred terms to help maintain standardized terms across studies, so
it is important for us to correct this before writing out the template.

`R2i::checkTemplate()` will throw an error for columns with non-matching controlled values.  However,
to see preferred value columns with issues, you must change the default `quiet` argument to `FALSE`
to receive messages.

In the case of **study**, we can see which columns have such terms by using the `getLookups()` 
function. This function's only argument is the ImmPort Template Name.

```{r get-lookups}
getIPLookups(ImmPortTemplateName = "study")
```

It appears that **study** has one column with controlled terms: *Age Unit*. To 
see what values are allowed for *Age Unit* we can use the `getLookupValues()` function. This
function takes in the ImmPort Template Name and the column name as arguments, then returns
a vector of allowed values.

```{r get-lookup-vals}
getIPLookupValues(ImmPortTemplateName = "study", templateColname = "Age Unit")
```

Since it looks like the original entry for **Age Unit** ('years') is not in the vector,
we must correct it with a capitalized version so it passes the `checkTemplate()` function
later on.  We can then see if our data frame passes our checks again.

```{r fix-years}
study$`Age Unit` <- "Years"

# NOTE: messages are printed out
results <- tryCatch(checkTemplate(df = study),
                    error = function(e){ return(e) })

# to see error statement
results
```

### PREPARING OTHER DATA FRAMES
Similar to **study**, the other 9 dataframes that make up the basicStudyDesign.txt can be accessed 
with the `getTemplateDF()` function.  Depending on the amount of information that needs to be
entered, it may be easier to use `edit()` or base R.  For the purpose of the vignette, an R-based
approach is used for reproducibiity and demonstration.

The next one is a simple one called **study_categorization** that defines the type of study performed
```{r study-categorization}
study_categorization <- getTemplateDF("study_categorization")
study_categorization[1,] <- c("Immune Response")
DT::datatable(study_categorization, options = list(scrollX = TRUE))
```

The next one is a simple one called **study_2_condition_or_disease** that defines the disease or condition studied.
```{r study-2-condtion-or-disease}
study_2_condition_or_disease <- getTemplateDF("study_2_condition_or_disease")
study_2_condition_or_disease[1,] <- c("Typhoid")
DT::datatable(study_2_condition_or_disease, options = list(scrollX = TRUE))
```

The next one we focus on is called **arm_or_cohort**. In this
demonstration case, we have a csv already of the information needed and just want to bind this
new information to correct headers.  Therefore we use the arm_or_cohort dataframe only for the
colnames() call.  An important note: the `checkTemplate()` function needs the "templateName"
attribute in the data frame in order to run the necessary checks.  This is easily done by
using `attr(arm_or_cohort, "templateName") <- "arm_or_cohort"`.

```{r make-arm-or-cohort-df}
arm_or_cohort <- getTemplateDF("arm_or_cohort")
aocImport <- read.table("arm_or_cohort_demo.tsv", sep = "\t", stringsAsFactors = FALSE)
colnames(aocImport) <- colnames(arm_or_cohort)
aocImport <- aocImport[ aocImport$`User Defined ID` != "", ]
DT::datatable(aocImport, options = list(scrollX = TRUE))

# to be consistent we rename aocImport for use in the 'write' functions later
arm_or_cohort <- aocImport

# Set "templateName" attribute to pass `checkTemplate()` fn
attr(arm_or_cohort, "templateName") <- "arm_or_cohort"
```

------

Making **study_personnel**:

```{r }
study_personnel <- getTemplateDF("study_personnel")
study_personnel[1,] <- c("Personnel1",
                       "Dr.",
                       "Khanna",
                       "Elizabeth",
                       "",
                       "Major University",
                       123,
                       "ekhanna@major.edu",
                       "PI",
                       "Principal Investigator",
                       "Major University")
DT::datatable(study_personnel, options = list(scrollX = TRUE))
```

------

Making **planned_visit**:

```{r }
planned_visit <- getTemplateDF("planned_visit")

planned_visit[1,] <- list(1,
                      "Screening",
                      1,
                      -10,
                      -2,
                      "",
                      "")

planned_visit[2,] <- list(2,
                      "Immunazation",
                      2,
                      0,
                      0,
                      "",
                      "")

planned_visit[3,] <- list(3,
                      "Chellenge",
                      3,
                      100,
                      110,
                      "",
                      "")

DT::datatable(planned_visit)
```

------

Making **inclusion_exclusion**: 

```{r make-inclusion-exclusion-df}
inclusion_exclusion <- getTemplateDF("inclusion_exclusion")
inclusion_exclusion[1,] <- c("InclExcl1",
                            "older than 35 years old",
                            "Exclusion")
DT::datatable(inclusion_exclusion)
```

------

**study_2_protocol** is different than other templates.  It is a small dataframe with only 1 row 
and two columns with the first column being a name and the second being he value.  

```{r make-study-2-protocol}
study_2_protocol <- getTemplateDF("study_2_protocol")
study_2_protocol[1,] <- "protocol 3445"
DT::datatable(study_2_protocol)
```

------

Making **study_file**:

```{r make-study-file}
study_file <- getTemplateDF("study_file")
study_file[1,] <- list("Appendix.txt",
                       "Study Appendix",
                       "Study Data")
DT::datatable(study_file)
```

------

Making **study_link**:

```{r make-study-link}
study_link <- getTemplateDF("study_link")
study_link[1,] <- c("main website",
                   "https://drkhannalab.major.edu/NewStudy1")
DT::datatable(study_link, options = list(autoWidth = TRUE))
```

------

**study_pubmed** is going to be left blank as a demonstration since some studies may need to be 
imported prior to being published.  Publication information can be entered later using an update
template.

```{r make-study-pubmed}
study_pubmed <- getTemplateDF("study_pubmed")
```

------

Before transforming our data frames and writing the tsv output, we can also do some quality assurance
checks of the text in our data frames using the text cleaning functions in the `R2i` package.

We will use `planned_visit$Name` as an example to first check for spelling errors using 
the `checkSpelling()` function that imports the `hunspell` package.

```{r finding-spelling-errors}
checkSpelling(input = planned_visit$Name)
```

If working interactively, you can use `interactiveReplace()` to go through each error found
with checkSpelling() and input a replacement at the prompt.  In this case, we will simply use
the `findReplace()` function to fix *Immunazation* and *Chellenge*.
```{r correcting-spelling-errors}
tmp <- findReplace(input = planned_visit$Name, find = "Immunazation", replace = "Immunization")
tmp <- findReplace(input = tmp, find = "Chellenge", replace = "Challenge")

planned_visit$Name <- tmp
DT::datatable(planned_visit)
```
------
### CREATING THE OUTPUT TSV
To create the tsv that will be included in the ImmPort submission, we use the
`transform_basicStudyDesign()` function that takes a named list of the 9 dataframes as the first
argument, as well as `outputDir` and `validate`.  The `outputDir` argument is the filepath for the
output directory where the tsv should be saved.  `validate` is a boolean with `TRUE` as the default
that uses the validator scripts from ImmPort's web application to ensure that the tsv meets the
criteria necessary for import.  To see more information about the `transform_basicStudyDesign()`
function you can always enter `?transform_basicStudyDesign` in the console.

```{r transform-data-into-tsv}
blocks <- list("study" = study,
               "study_categorization" = study_categorization,
               "study_2_condition_or_disease" = study_2_condition_or_disease,
               "arm_or_cohort" = arm_or_cohort,
               "study_personnel" = study_personnel,
               "planned_visit" = planned_visit,
               "inclusion_exclusion" = inclusion_exclusion,
               "study_2_protocol" = study_2_protocol,
               "study_file" = study_file,
               "study_link" = study_link,
               "study_pubmed" = study_pubmed)

transform_basicStudyDesign(blocks = blocks,
                           outputDir = NULL,
                           validate = TRUE)
```



